Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-09-10 documents/ovando-etal-assessing-global-fisheries.Rmd
jims edits

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess. Our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
---

@@

XX discussion with Jim. How can we frame this. I would actually push back the idea that we have learned about stock status. I buy that the catch data can update your prior on growth parameters, but by the rules we're playing with here, it shouldn't update our prior on stock status should it. Now, the only thing I can't wrap my head aroudn is the case where catches are all zero, and you have a prior that the stock is collapsed. I guess my question is from a strikely Bayes theorem perspective, where in here is your prior on stock status being updated, since there's no likelihood as written, except for the binomial process that says that the stock exists.

For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores. We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock. We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so. We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status. We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`. It is critical to note that in both of these cases the model has learned nothing new about recent stock status that was not encoded in the supplied priors, catch data, and functional form of the model; `sraplus` simply helps reveal our implicit priors on other parameters (e.g. maximum sustainable yield, MSY), and provides timeseries of measures such as B/B~MSY~. That, in the absence of data to actually fit to such as an index of abundance, methods such as CMSY or sraplus can learn anything new about a stock's status that is not already implicitly coded in the priors, catch data, or specified functional relationships, is a common misconception (see [Methods](#Methods) a for detailed explanation of why this is). Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.

@@

Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world. While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models using each of the available datasets beside the RAM-derived indices produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).

@@

First, use of effort data requires making an assumption about the rate at which the fishing power of a unit of fishing effort has changed over time. @rousseau2019 used a fixed global rate of 2.6% to describe the increasing trend in fishing power, but as shown in their paper and in our analysis, any estimates of stock status using these effort data will be critically sensitive to the value of this parameter. Global studies of technology creep could help inform more realistic values and functional forms for the evolution of fishing skill in particular fisheries and regions, and also require accounting for regional complications including density-dependent catchability [@thorson2010]. Second, the effort reconstructions as employed here are still at very coarse scales, requiring us to apply one effort series for all the stocks within a given region. Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes. Along with supporting methods used here, access to fine-scale effort dynamics data can help inform data-limited assessment based on catch or length-composition data XX @thorson2013;@ovando2018XX

@@

All of our estimates are Bayesian in nature. We can break the use of `sraplus` into two distinct categories: with data and without. By "data", we refer to measurements which are used to confront model estimates within a likelihood function. In our context, these include fishery-independent survey data, or a CPUE index. When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters). Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e. results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches. This step updates the prior distribution XX of population parameters XX by eliminating combinations of priors that are impossible for a given catch history.

The full list of estimable parameters are listed in Table.\@ref(tab:param-tab). *r* and *k* are the only two parameters that are always estimated. Estimation of every other parameter can be turned on or off. When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user. For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *k*, *m*, $\sigma_{obs}$, $\gamma$, and *B0*. *q* is also estimated when needed.

[@thorson2017c]

@@

`sraplus` can be run in two forms: either as a stock reduction analysis [SRA,@walters2006], or fit to an index of abundance (fishery dependent or independent). Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. `sraplus` allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.

It is critical to note that a stock reduction analysis is only a mechanism for revealing implicit priors on fishery status and parameters through a sample-importance-resampling algorithm: the model cannot learn anything new about stock status in the absence of additional data to fit to, beyond the prior on stock status that is fed to it and the constraints imposed by structural assumptions. The usefulness of this process then is in filling in the gap of parameters of interest (e.g. maximum sustainable yield), and providing historic values of stock status. Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" is estimated through stock reduction analysis. All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.

@@

Priors for all estimable parameters can be left at default values or set by the user. The shape parameter is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimate the ratio of B~MSY~ to K for many fish taxa. While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on @thorson2012 for on the genus of the species in question.

@@

Suppose that the only thing we observe from a fishery is a catch history. Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past. We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.  Beyond that though, baring additional information we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0.

@@

We present an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000]. Our solution amounts to a two-step SIR algorithm. We first run the standard SIR algorithm as described above. We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.

@@

And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$. The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status. In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status". This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot)).

@@

biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
                                      estimate_shape = FALSE)

@@

co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
                                      estimate_shape = FALSE)

@@

```{r}
plot_prior_posterior(co_fit, co_driors)
```

>>>>

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess. Obtaining accurate estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of high-priority data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
---

@@

For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores. We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock. We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so. We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status. We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`. It is critical to note that in both of these cases the catch data on its own contains no information about stock status; we can only produce estimates of stock status by combining catch data with explicit priors and specified functional relationships [@thorson2013]. This is important since it means that we must look to our priors and functional forms to understand where our insight stock status is coming from, not the catch data alone. Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.

@@

Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world. While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).

@@

First, use of effort data requires making an assumption about the rate at which the fishing power of a unit of fishing effort has changed over time. @rousseau2019 used a fixed global rate of 2.6% to describe the increasing trend in fishing power, but as shown in their paper and in our analysis, any estimates of stock status using these effort data will be critically sensitive to the value of this parameter. Global studies of technology creep could help inform more realistic values and functional forms for the evolution of fishing skill in particular fisheries and regions, and also require accounting for regional complications including density-dependent catchability [@thorson2010]. Second, the effort reconstructions as employed here are still at very coarse scales, requiring us to apply one effort series for all the stocks within a given region. Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes. Along with supporting methods used here, access to fine-scale effort dynamics data can help inform data-limited assessment based on catch or length-composition data [@ovando2018; @thorson2013].

@@

All of our estimates are Bayesian in nature. We can break the use of `sraplus` into two distinct categories: with data and without. By "data", we refer to measurements which are used to confront model estimates within a likelihood function. In our context, these include fishery-independent survey data, or a CPUE index. When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters). Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e. results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches. This step updates the prior distribution of population parameters by eliminating combinations of priors that are impossible for a given catch history and a specified functional form.

The full list of estimable parameters are listed in Table.\@ref(tab:param-tab). *r* and *k* are the only two parameters that are always estimated. Estimation of every other parameter can be turned on or off. When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user. For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *k*, $\sigma_{obs}$, $\gamma$, and *B0*. *q* is also estimated when needed.

@@

`sraplus` can be run in two forms: either as a stock reduction analysis [SRA,@walters2006], or fit to an index of abundance (fishery dependent or independent). Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. `sraplus` allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis. All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.

@@

Priors for all estimable parameters can be left at default values or set by the user. The shape parameter is usually not reliably estimable given available data for surplus production models, however, Thorson et al. 2012 [@thorson2012] provides estimate the ratio of B~MSY~ to K for many fish taxa. While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on Thorson et al. 2012 [@thorson2012] for the genus of the species in question.

@@

Suppose that the only thing we observe from a fishery is a catch history. Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past. We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.  Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0.

@@

We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000]. Our solution amounts to a two-step SIR algorithm. We first run the standard SIR algorithm as described above. We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.

@@

And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$. The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status. In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status". This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).

@@

biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
                                      estimate_shape = FALSE, estimate_proc_error = TRUE)

@@

co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
                                      estimate_shape = FALSE, 
                               estimate_proc_error = TRUE)

@@

```{r ppplot, fig.cap="Prior-posterior plots of fits for case study fishery in Fig.6"}

a = get_prior_posterior(co_fit, co_driors)

a$prior_posterior %>% 
  ggplot(aes(x = variable, y = mean, ymin = lower, ymax = upper, color = source)) + 
  geom_pointrange(position = position_dodge(width = 0.1)) + 
  facet_wrap(~variable, scales = "free") + 
  theme(axis.text.x = element_blank()) + 
  scale_color_discrete(name = '') + 
  scale_y_continuous(name = "Value") + 
  scale_x_discrete(name = '') + 
  theme_minimal()
